\section*{Discussion}

The results of benchmark we showed a huge increase in performance with Numba,
Cython and Armadillo c++. Performing between 100 to 500 times faster than the standard python implementation.
For n less than 100 the Numba version were the fastest being around 20
\% faster than Cython and around 50 \% faster than Armadillo (\cref{fig:comp_python}). This may be due to the
efficiency of machine code compiled by Numba jit. For large n the Cython version
showed the best performance being around five times faster than Armadillo. The
Numba version was around 3.5 times faster than Armadillo (\cref{fig:comp_c++}).
The Numba version did stay constant at around 3.5 times the speed of Armadillo.
While Cython version show a steady increase in relative speed up of around from
3 for n = 100 to 5.5 for n = 350. 


Surprising to us was that both the Numba and Cython would beat Armadillo by such
a large margin. The performance of Numba was especially impressive considering
that only change we did compared to the baseline python code was adding the
@jit(nopython =True) deceleration at the top of our functions. This results may
be because of the high level features that Armadillo offers does compromise some
performance. This something that would be interesting to investigate further by
comparing an pure C++ implementation with the Armadillo implementation.

We did not only evaluate the performance of each version, preserving code
readability, ease of use and robustness where also considered.  

