\section*{Introduction}	

 High-level programming languages have in the recent years become increasingly	
popular in scientific programming. The high-level syntax renders the code very	
readable,	
allowing the code for instance to closely resemble mathematics. The nice syntax	
combined with excellent and easy to use libraries	
for plotting and mathematics makes Python or Matlab the	
programming languages of choice for many scientists.	
The nice high-level features does not come without penalty, for	
instance running large loops in Python or Matlab is known to be notoriously slow.	
Therefore in high performance computing languages low level languages like C++/C	
or Fortran are still the standard.	

 Fortunately there exists solutions which tries to offer both the readable syntax	
and ease of use of a high-level language and performance of low level languages.	
We will	
look at three different solutions, the just in time (JIT) compiler for python	
Numba, Cython which is a C based extension of python and Armadillo a	
library for scientific computing a in c++ offering a more high level syntax.	

 First we will give an short description of Numba, Cython and Armadillo.	
Next we will describe our experimental setup, where we implement the Jacobi	
method for finding eigenvalues using the aforementioned approaches and	
benchmarking them against a standard python approach. The specific problem we	
will solve is the wave equation for a buckling beam.	


 \subsection*{Numba}	
Numba is an open source just in time (JIT) compiler which translates a subset of	
python code	
into fast machine code. Currently Numba works best with NumPy, functions	
and loops. Enabling numba in a python script is very simple and is done by	
simply typing in a @jit declaration above the python function you want to speed	
up. If the @jit(nopython = true) Numba will compile the entire python function	
without involving python at all, resulting in maximum speed up. Still to gain	
any increase in speed at all, the code needs to consist of	
numpy functions or loops.	

 \subsection*{Cython}	
Cython is an optimising static compiler for Cython, which is a superset of the	
python programming language. The cython compiler translate Python into C code by	
adding static type decelerations. Cython offers better readability of high-level	
python code combined with the speed of low level C and C++. This makes Cython ideal for	
creating fast modules which can speed up otherwise slow python code. One of the	
great advantages with Cython when developing code compared to for instance c++	
or C, is that since Cython is a superset of Python we can actually start with	
our slow Python code and then optimize the code iteratively. Compare this	
approach to translating an entire python program into c++/c, which will probably	
lead to mistakes during translation and unnecessary time spent debugging.	

 \subsection*{Armadillo}	
Armadillo is an open-source linear algebra library for c++. Armadillo aims to	
be efficient while at the same time providing an easy-to-use interface. Armadillo	
has a large variety of different linear algebra functions provided through	
integration of the linear algebra package (LAPACK).	
