\section{Introduction}
Rossby waves was first introduced by the Swedish meteorologist
Carl G Rossby in 1939 \parencite{Rossby1939}. Rossby waves are large scale
planetary waves, with typical scales ranging from 
hundreds to thousands of kilometres. They play a major role in dictating
the weather on daily and large time scales and are crucial for the meridional
transport of heat, moisture and momentum \parencite{midSynDyn}. The rossby wave phenomenon arise due to the
latitudinal variation in the Coriolis acceleration and are observed both in
ocean an atmosphere. 

In this study we will examine Rossby waves in a simplified 
atmosphere/ocean. We will assume that our atmosphere/ocean is barotropic, that
is the density depends only on pressure, which via the ideal gas law holds that
the temperature is constant on a isobaric surface, i.e. no stratification. We also assume the flow to be frictionless and
non-divergent, consequently neglecting vertical velocities. Under these
simplifications the quasi geostrophic vorticity equation is reduced to
\cref{eq:QG_vorticity}, where $\zeta$ is the relative vorticity and $f$ is the
Coriolis parameter, $f=2\Omega \sin(\theta)$, where $\Omega$ is
the rotation rate of the earth $\Omega = 2\pi/ \mathrm{day}$.
\begin{equation}\label{eq:QG_vorticity}
    \frac{D(\zeta + f)}{Dt} = 0
\end{equation}
In the simplified the quasi geostrophic vorticity equation the
change in absolute vorticity $(\zeta + f)$ following the flow is zero, in other
words the absolute vorticity is conserved. When you have the vorticity in terms
of a conserved quantity it is commonly called potential vorticity (PV). If a fluid parcel is moving
poleward towards region of larger $f$ then, there must be a compensating change
in $\zeta$ in order for the absolute vorticity to remain unchanged. This
balancing act between planetary and relative vorticity is the essential
mechanism behind
Rossby waves. We will derive the barotropic Rossby
wave equation (BRWE) from \cref{eq:QG_vorticity} in the subsequent section.

After we have developed our analytical framework, we will discretize the BRWE
and solve the BRWE in both one and two dimension, using both periodic and
bounded boundary condition. We will examine both the forward
difference and central difference scheme and consider their stability. 